---
title: "Nature-based Solutions and Political Risk: A Cross-Country Analysis"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
    fig_width: 10
    fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
# Set working directory to project root (one level above analysis/)
knitr::opts_knit$set(root.dir = normalizePath(file.path(knitr::current_input(dir = TRUE), "..")))
```

## Overview

This document examines how country-level **total Nature-based Solutions (NbS)
mitigation potential** correlates with **World Governance Indicators (WGI)** and
whether political risk shapes **market-based NbS implementation**. Three datasets
are integrated:

1. **Naturebase v3** (2025) — biophysical NbS potential in tCO₂e/yr for 20
   pathways across 216 countries (`total_ncs_tco2e`)
2. **World Bank WGI 2024** — six governance dimensions across 216 countries
3. **Karnik et al. 2024** (*Zenodo* doi:10.5281/zenodo.11459391) — 575
   market-based carbon offset projects (AD, IFM, ARR) in 55 countries

**Political risk dimensions (WGI):**

| Code      | Dimension                                      |
|-----------|------------------------------------------------|
| CC.EST    | Control of Corruption                          |
| GE.EST    | Government Effectiveness                       |
| PV.EST    | Political Stability & Absence of Violence      |
| RL.EST    | Rule of Law                                    |
| RQ.EST    | Regulatory Quality                             |
| VA.EST    | Voice & Accountability                         |

> **Additional source — Fragile States Index (FSI):** Download the latest annual
> CSV from <https://fragilestatesindex.org/> and place it in
> `data/fsi/fsi_latest.csv` to activate Section 8. The FSI provides 12
> sub-indicators (Security Apparatus, State Legitimacy, Factionalized Elites,
> etc.) that complement the broader WGI governance dimensions.

---

## 1. Package Setup

```{r packages}
# Install any missing packages before running:
# install.packages(c("tidyverse", "readxl", "ggcorrplot", "ggrepel",
#                    "countrycode", "patchwork", "broom", "scales", "here"))

library(tidyverse)
library(readxl)         # Read local WGI Excel files
library(ggcorrplot)     # Correlation heatmaps
library(ggrepel)        # Non-overlapping labels on scatter plots
library(countrycode)    # Country code harmonisation
library(patchwork)      # Combine ggplots
library(broom)          # Tidy model output
library(scales)
library(here)
```

---

## 2. Load Naturebase Data

```{r load-nbs}
nbs_raw <- read_csv(here("data/all_pathways/all_pathways_tco2e_adm0.csv"))

# ── Pathway display labels ─────────────────────────────────────────────────────
pathway_labels <- c(
  # Individual pathways
  crp_cba_tco2e  = "Cropland Agroforestry",
  crp_rec_tco2e  = "Reduced Emissions (Crops)",
  crp_scc_tco2e  = "Soil Carbon (Crops)",
  for_afc_tco2e  = "Avoided Forest Conversion",
  for_csf_tco2e  = "Climate-Smart Forestry",
  for_ref_tco2e  = "Forest Reforestation",
  for_rwf_tco2e  = "Reduced Woodland Fire (Forest)",
  grs_agc_tco2e  = "Agroforestry (Grassland)",
  grs_asc_tco2e  = "Avoided Savanna Conversion",
  grs_grr_tco2e  = "Grassland Restoration",
  grs_reg_tco2e  = "Reforestation (Grassland)",
  grs_rwf_tco2e  = "Reduced Grassland Fire",
  grs_sba_tco2e  = "Savanna Burning Management",
  grs_scg_tco2e  = "Soil Carbon (Grassland)",
  grs_sfm_tco2e  = "Sustainable Forest Mgmt",
  wet_apc_tco2e  = "Avoided Peatland Conversion",
  wet_awc_tco2e  = "Avoided Wetland Conversion",
  wet_cwr_tco2e  = "Coastal Wetland Restoration",
  wet_ipm_tco2e  = "Improved Peatland Mgmt",
  wet_per_tco2e  = "Peatland Restoration",
  # Biome aggregates
  total_crop_tco2e    = "Total: Cropland",
  total_forest_tco2e  = "Total: Forest",
  total_grass_tco2e   = "Total: Grassland",
  total_wet_tco2e     = "Total: Wetland",
  total_peat_tco2e    = "Total: Peatland",
  # Activity-type aggregates
  total_protect_tco2e = "Total: Protection",
  total_manage_tco2e  = "Total: Management",
  total_restore_tco2e = "Total: Restoration",
  total_ncs_tco2e     = "Total NCS"
)

nbs_individual <- names(pathway_labels)[1:20]
nbs_biome      <- names(pathway_labels)[21:25]
nbs_activity   <- names(pathway_labels)[26:29]
nbs_grand      <- "total_ncs_tco2e"

cat("Countries in Naturebase:", nrow(nbs_raw), "\n")
glimpse(nbs_raw)
```

---

## 3. Load World Governance Indicators (Local Excel)

Data are read from the locally downloaded WGI 2025 dataset
(`data/wgidataset_with_sourcedata-2025.xlsx`). The most recent year (2024)
is used for each country.

```{r load-wgi}
# Short alias → display label (used for plot titles, table headers)
wgi_codes <- c(
  CC.EST = "Control of Corruption",
  GE.EST = "Government Effectiveness",
  PV.EST = "Political Stability & No Violence",
  RL.EST = "Rule of Law",
  RQ.EST = "Regulatory Quality",
  VA.EST = "Voice & Accountability"
)

# Sheet name in the Excel file for each dimension
wgi_sheet_map <- c(
  CC.EST = "cc",
  GE.EST = "ge",
  PV.EST = "pv",
  RL.EST = "rl",
  RQ.EST = "rq",
  VA.EST = "va"
)

wgi_excel_path <- here("data", "wgidataset_with_sourcedata-2025.xlsx")

# Read each sheet and extract the governance estimate for the most recent year
read_wgi_sheet <- function(dim_code) {
  read_excel(wgi_excel_path, sheet = wgi_sheet_map[dim_code]) %>%
    rename(
      wb_code      = `Economy (code)`,
      country      = `Economy (name)`,
      region       = Region,
      income_level = `Income classification`,
      year         = Year,
      estimate     = `Governance estimate (approx. -2.5 to +2.5)`
    ) %>%
    select(wb_code, country, region, income_level, year, estimate) %>%
    filter(!is.na(estimate)) %>%
    group_by(wb_code) %>%
    slice_max(year, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    rename(!!dim_code := estimate)
}

wgi_sheets <- lapply(names(wgi_sheet_map), read_wgi_sheet)

# Wide format: one row per country, one column per WGI dimension
wgi_wide <- wgi_sheets[[1]] %>%
  select(wb_code, country, region, income_level, CC.EST)
for (i in 2:length(wgi_sheets)) {
  dim_col <- names(wgi_sheet_map)[i]
  wgi_wide <- left_join(
    wgi_wide,
    wgi_sheets[[i]] %>% select(wb_code, !!sym(dim_col)),
    by = "wb_code"
  )
}

# Convert World Bank country codes → ISO3
# countrycode handles most; manual overrides for legacy WB codes
wgi <- wgi_wide %>%
  mutate(
    iso3c = countrycode(wb_code, origin = "wb", destination = "iso3c",
                        warn = FALSE),
    iso3c = case_when(
      wb_code == "ADO" ~ "AND",   # Andorra
      wb_code == "ZAR" ~ "COD",   # DR Congo
      wb_code == "ROM" ~ "ROU",   # Romania
      wb_code == "TMP" ~ "TLS",   # Timor-Leste
      wb_code == "PSE" ~ "PSE",   # Palestine
      !is.na(iso3c)   ~ iso3c,
      TRUE             ~ wb_code
    ),
    income_level = factor(income_level,
      levels = c("Low income", "Lower middle income",
                 "Upper middle income", "High income"))
  )

cat("WGI records loaded:", nrow(wgi), "\n")
```

---

## 4. Merge NbS + Governance Data

```{r merge}
df <- nbs_raw %>%
  rename(iso3c = adm0_id) %>%
  left_join(wgi, by = "iso3c") %>%
  filter(!is.na(CC.EST))  # require at least one WGI estimate

cat("Countries matched:", nrow(df), "\n")
cat("Countries without WGI (dropped):", nrow(nbs_raw) - nrow(df), "\n")

# Compute per-hectare NCS density for size-normalised comparisons
df <- df %>%
  mutate(
    across(
      all_of(c(nbs_individual, nbs_biome, nbs_activity, nbs_grand)),
      ~ . / adm0_ha,
      .names = "{.col}_dens"   # tCO2e / ha_country / yr
    )
  )
```

---

## 5. Descriptive Statistics

```{r descriptives}
desc <- df %>%
  select(total_ncs_tco2e, all_of(names(wgi_codes))) %>%
  pivot_longer(everything()) %>%
  group_by(name) %>%
  summarise(
    n      = sum(!is.na(value)),
    mean   = mean(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    sd     = sd(value, na.rm = TRUE),
    min    = min(value, na.rm = TRUE),
    max    = max(value, na.rm = TRUE)
  ) %>%
  mutate(
    label = coalesce(pathway_labels[name], wgi_codes[name], name)
  ) %>%
  relocate(label, .after = name)

knitr::kable(desc, digits = 2,
  caption = "Descriptive statistics — Total NCS potential (tCO₂e/yr) and WGI scores")
```

---

## 6. Total NCS × WGI Spearman Correlations

Spearman rank correlations between **log1p(Total NCS tCO₂e/yr)** and each
WGI governance dimension. p-values are from a two-sided test with a
continuity correction.

```{r cor-total-ncs}
wgi_vars <- names(wgi_codes)

# Compute Spearman rho + p-value for each WGI dim vs total NCS
cor_ncs <- map_dfr(wgi_vars, function(wv) {
  tmp <- df %>%
    filter(!is.na(total_ncs_tco2e), total_ncs_tco2e > 0, !is.na(.data[[wv]])) %>%
    mutate(log_ncs = log1p(total_ncs_tco2e))
  ct  <- cor.test(tmp$log_ncs, tmp[[wv]], method = "spearman", exact = FALSE)
  tibble(
    wgi_var   = wv,
    wgi_label = wgi_codes[wv],
    rho       = ct$estimate,
    p_value   = ct$p.value,
    n         = nrow(tmp),
    sig       = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      TRUE            ~ ""
    )
  )
})

cor_ncs %>%
  select(wgi_label, rho, p_value, n, sig) %>%
  mutate(across(c(rho, p_value), ~ round(.x, 4))) %>%
  knitr::kable(
    col.names = c("WGI Dimension", "Spearman ρ", "p-value", "N", "Sig."),
    caption   = "Spearman correlations: log1p(Total NCS tCO₂e/yr) × WGI (2024)"
  )
```

```{r cor-bar, fig.height=4, fig.width=7}
cor_ncs %>%
  ggplot(aes(x = rho, y = reorder(wgi_label, rho),
             fill = rho > 0)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = paste0(round(rho, 3), sig),
                x     = rho + ifelse(rho >= 0, 0.01, -0.01)),
            hjust = ifelse(cor_ncs$rho >= 0, 0, 1), size = 3.8) +
  geom_vline(xintercept = 0, linewidth = 0.5) +
  scale_fill_manual(values = c("TRUE" = "#1a9850", "FALSE" = "#d73027")) +
  scale_x_continuous(limits = c(-0.4, 0.6), expand = expansion(mult = 0.02)) +
  labs(
    title    = "Spearman ρ: Total NCS Potential × WGI Governance Dimensions",
    subtitle = "* p<0.05  ** p<0.01  *** p<0.001  |  log1p(tCO₂e/yr), N≈200 countries",
    x = "Spearman ρ", y = NULL
  )
```

---

## 7. Scatter Plots: Total NCS vs. Each WGI Dimension

Each panel shows **log₁₀(Total NCS tCO₂e/yr)** against one WGI dimension.
Countries in the top 10% of NCS potential or bottom 10% of governance are
labelled. The OLS trend line (with 95% CI) is shown in black.

```{r scatter-all-wgi, fig.height=9, fig.width=11}
label_thresh <- quantile(df$total_ncs_tco2e, 0.90, na.rm = TRUE)

make_scatter <- function(wv) {
  risk_thresh <- quantile(df[[wv]], 0.10, na.rm = TRUE)
  df %>%
    filter(!is.na(.data[[wv]]), total_ncs_tco2e > 0) %>%
    mutate(
      log_ncs = log10(total_ncs_tco2e),
      label   = if_else(
        total_ncs_tco2e > label_thresh | .data[[wv]] < risk_thresh,
        adm0_name, NA_character_
      )
    ) %>%
    ggplot(aes(x = .data[[wv]], y = log_ncs)) +
    geom_point(aes(color = income_level), alpha = 0.70, size = 1.9) +
    geom_smooth(method = "lm", se = TRUE, color = "black",
                linewidth = 0.7, fill = "grey80") +
    geom_text_repel(aes(label = label), size = 2.4, max.overlaps = 12,
                    segment.color = "grey55", segment.size = 0.3) +
    scale_color_brewer(palette = "RdYlBu", name = NULL, na.value = "grey70",
                       guide = guide_legend(nrow = 2)) +
    labs(
      x     = wgi_codes[wv],
      y     = "log₁₀(Total NCS tCO₂e/yr)",
      title = wgi_codes[wv]
    ) +
    theme_minimal(base_size = 10) +
    theme(legend.position = "bottom",
          legend.text     = element_text(size = 7),
          plot.title      = element_text(face = "bold", size = 10))
}

plots <- map(wgi_vars, make_scatter)

wrap_plots(plots, ncol = 3, guides = "collect") &
  theme(legend.position = "bottom")
```

---

## 8. Fragile States Index (Optional)

Place the annual FSI CSV (`fsi_latest.csv`) in `data/fsi/` to run this section.
Expected columns: `Country`, `Total`, plus 12 sub-indicator columns
(C1–C3, S1–S2, E1–E3, P1–P3, X1, and/or their full names).

```{r fsi, eval=file.exists(here("data/fsi/fsi_latest.csv"))}
fsi_raw <- read_csv(here("data/fsi/fsi_latest.csv"))

# Harmonise country names → ISO3 via countrycode
fsi <- fsi_raw %>%
  mutate(
    iso3c = countrycode(Country, origin = "country.name", destination = "iso3c",
                        warn = FALSE)
  ) %>%
  filter(!is.na(iso3c)) %>%
  rename(fsi_total = Total) %>%
  select(iso3c, fsi_total, any_of(c(
    "C1: Security Apparatus", "C2: Factionalized Elites", "C3: Group Grievance",
    "E1: Economy", "E2: Economic Inequality", "E3: Human Flight and Brain Drain",
    "P1: State Legitimacy", "P2: Public Services", "P3: Human Rights",
    "S1: Demographic Pressures", "S2: Refugees and IDPs", "X1: External Intervention"
  )))

df_fsi <- df %>%
  left_join(fsi, by = "iso3c")

fsi_vars <- names(fsi)[-1]

cor_fsi_mat <- df_fsi %>%
  select(all_of(c(nbs_biome, nbs_grand, fsi_vars))) %>%
  mutate(across(all_of(c(nbs_biome, nbs_grand)), log1p)) %>%
  cor(method = "spearman", use = "pairwise.complete.obs")

cor_fsi_sub <- cor_fsi_mat[c(nbs_biome, nbs_grand), fsi_vars]
rownames(cor_fsi_sub) <- pathway_labels[rownames(cor_fsi_sub)]

knitr::kable(round(cor_fsi_sub, 3),
             caption = "Spearman rho — NbS biome totals × FSI sub-indicators")
```

---

## 9. Detailed Scatter: Total NCS vs. Political Stability

High-resolution view of the strongest governance relationship, with all
countries labelled by income group and region.

```{r scatter-ps-detail, fig.height=7, fig.width=10}
df %>%
  filter(!is.na(PV.EST), total_ncs_tco2e > 0) %>%
  mutate(
    log_ncs = log10(total_ncs_tco2e),
    label   = if_else(
      total_ncs_tco2e > quantile(total_ncs_tco2e, 0.85, na.rm = TRUE) |
      PV.EST < quantile(PV.EST, 0.10, na.rm = TRUE),
      adm0_name, NA_character_
    )
  ) %>%
  ggplot(aes(x = PV.EST, y = log_ncs)) +
  geom_point(aes(color = income_level, shape = region), alpha = 0.80, size = 2.8) +
  geom_smooth(method = "lm", se = TRUE, color = "black",
              linewidth = 0.9, fill = "grey85") +
  geom_text_repel(
    aes(label = label), size = 2.8, max.overlaps = 25,
    segment.color = "grey55", segment.size = 0.3,
    box.padding = 0.4
  ) +
  scale_color_brewer(palette = "RdYlBu", name = "Income Group",
                     na.value = "grey60") +
  scale_shape_manual(
    name   = "Region",
    values = setNames(c(16, 17, 15, 18, 8, 4, 3),
                      na.omit(unique(df$region)))
  ) +
  labs(
    x       = "Political Stability & Absence of Violence (WGI 2024, −2.5 to +2.5)",
    y       = "Total NCS Potential (log₁₀ tCO₂e/yr)",
    title   = "Total NCS Potential vs. Political Stability",
    caption = "Naturebase v3 (2025); World Governance Indicators 2024"
  ) +
  theme_minimal(base_size = 12) +
  guides(shape = guide_legend(ncol = 1))
```

---

## 10. Sub-group Analysis by Income Group

Stratified Spearman correlations between Total NCS and each WGI dimension,
computed separately for each World Bank income group.

```{r subgroup-income-cor}
income_cors <- df %>%
  filter(!is.na(income_level)) %>%
  group_by(income_level) %>%
  summarise(
    n = n(),
    across(
      all_of(wgi_vars),
      ~ cor(log1p(total_ncs_tco2e), .x,
            method = "spearman", use = "complete.obs"),
      .names = "{.col}"
    )
  ) %>%
  pivot_longer(all_of(wgi_vars), names_to = "wgi_var", values_to = "rho") %>%
  mutate(wgi_label = wgi_codes[wgi_var])

ggplot(income_cors,
       aes(x = rho, y = reorder(income_level, rho, FUN = mean),
           color = wgi_label)) +
  geom_point(size = 3.5) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_brewer(palette = "Dark2", name = "WGI Dimension") +
  labs(
    x     = "Spearman rho (log1p Total NCS ~ WGI)",
    y     = NULL,
    title = "NCS–Governance Correlations by Income Group"
  ) +
  theme_minimal(base_size = 12)
```

```{r subgroup-income-facet, fig.height=8, fig.width=11}
df %>%
  filter(!is.na(income_level), total_ncs_tco2e > 0) %>%
  pivot_longer(cols = all_of(wgi_vars),
               names_to = "wgi_var", values_to = "wgi_val") %>%
  mutate(wgi_label = wgi_codes[wgi_var]) %>%
  ggplot(aes(x = wgi_val, y = log1p(total_ncs_tco2e), color = income_level)) +
  geom_point(alpha = 0.4, size = 1.2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.9) +
  facet_grid(income_level ~ wgi_label,
             labeller = labeller(wgi_label = label_wrap_gen(12))) +
  scale_color_brewer(palette = "RdYlBu") +
  labs(
    x     = "WGI Score",
    y     = "Total NCS (log1p tCO₂e/yr)",
    title = "Total NCS ~ WGI, stratified by income group"
  ) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "none",
        strip.text = element_text(size = 7))
```

---

## 11. Sub-group Analysis by World Bank Region

```{r subgroup-region, fig.height=6, fig.width=12}
region_cors <- df %>%
  filter(!is.na(region)) %>%
  group_by(region) %>%
  summarise(
    n = n(),
    across(
      all_of(wgi_vars),
      ~ cor(log1p(total_ncs_tco2e), .x,
            method = "spearman", use = "complete.obs"),
      .names = "{.col}"
    )
  ) %>%
  pivot_longer(all_of(wgi_vars), names_to = "wgi_var", values_to = "rho") %>%
  mutate(wgi_label = wgi_codes[wgi_var])

ggplot(region_cors,
       aes(x = rho, y = reorder(region, rho, FUN = mean),
           fill = wgi_label)) +
  geom_col(position = position_dodge(0.85), width = 0.8) +
  geom_vline(xintercept = 0, linewidth = 0.5) +
  scale_fill_brewer(palette = "Set2", name = "WGI Dimension") +
  labs(
    x     = "Spearman rho (log1p Total NCS ~ WGI)",
    y     = NULL,
    title = "Regional Variation in NCS–Governance Correlations"
  ) +
  theme_minimal(base_size = 11)
```

---

## 12. OLS Regression: Total NCS ~ All WGI Dimensions

A single OLS model regresses log1p Total NCS on all six WGI dimensions
simultaneously. This estimates the partial association of each governance
dimension conditional on the others.

```{r ols-regression, fig.height=5, fig.width=8}
ols <- lm(
  log1p(total_ncs_tco2e) ~
    CC.EST + GE.EST + PV.EST + RL.EST + RQ.EST + VA.EST,
  data = df
)

summary(ols)

tidy(ols, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(term = wgi_codes[term]) %>%
  ggplot(aes(x = estimate,
             xmin = conf.low, xmax = conf.high,
             y = reorder(term, estimate))) +
  geom_pointrange(size = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    x        = "OLS Coefficient (95% CI)",
    y        = NULL,
    title    = "OLS: log1p(Total NCS) ~ WGI",
    subtitle = paste0("Adj. R² = ",
                      round(summary(ols)$adj.r.squared, 3),
                      " | N = ", nobs(ols))
  ) +
  theme_minimal(base_size = 12)
```

### Per-biome regressions

```{r ols-per-biome}
biome_models <- map(
  set_names(c(nbs_biome, nbs_grand)),
  ~ lm(
      reformulate(wgi_vars, response = paste0("log1p(", .x, ")")),
      data = df
    ) %>%
    tidy(conf.int = TRUE) %>%
    filter(term != "(Intercept)") %>%
    mutate(
      outcome   = pathway_labels[.x],
      wgi_label = wgi_codes[term]
    )
) %>%
  bind_rows()

ggplot(biome_models,
       aes(x = estimate, xmin = conf.low, xmax = conf.high,
           y = reorder(wgi_label, estimate),
           color = outcome)) +
  geom_pointrange(position = position_dodge(0.7), size = 0.55) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_brewer(palette = "Set1", name = "NbS Outcome") +
  facet_wrap(~outcome, nrow = 2) +
  labs(
    x     = "OLS Coefficient (95% CI)",
    y     = NULL,
    title = "Per-biome OLS: log1p(NbS potential) ~ WGI"
  ) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "none",
        axis.text.y = element_text(size = 8))
```

---

## 13. Market-based NbS: Karnik et al. (2024) Carbon Offset Projects

A global database of **575 nature-based carbon offset project boundaries** across
55 countries (Karnik et al., 2024, *Zenodo* doi:10.5281/zenodo.11459391).
Projects span three types: Avoided Deforestation (AD), Improved Forest Management
(IFM), and Afforestation / Reforestation / Revegetation (ARR). This represents
**market-based, implemented** NbS activity — a complement to Naturebase's
biophysical potential estimates.

```{r load-karnik}
karnik_raw <- read_csv(here("data/karnik2024/carbon_projects_database_index.csv"),
                       show_col_types = FALSE)

# Aggregate to country level
karnik_cty <- karnik_raw %>%
  mutate(
    iso3c = countrycode(`Country`, origin = "country.name",
                        destination = "iso3c", warn = FALSE)
  ) %>%
  filter(!is.na(iso3c)) %>%
  group_by(iso3c) %>%
  summarise(
    n_projects = n(),
    n_ad       = sum(`Project Type` == "AD",  na.rm = TRUE),
    n_ifm      = sum(`Project Type` == "IFM", na.rm = TRUE),
    n_arr      = sum(`Project Type` == "ARR", na.rm = TRUE),
    n_verra    = sum(`Registry Name` == "Verra", na.rm = TRUE),
    .groups = "drop"
  )

cat(sprintf("Karnik: %d countries with ≥1 project | %d projects total\n",
            nrow(karnik_cty), sum(karnik_cty$n_projects)))
```

### 13.1 Project Count by Country

```{r karnik-bar, fig.height=6}
karnik_cty %>%
  left_join(select(df, iso3c, country_name = adm0_name, region), by = "iso3c") %>%
  arrange(desc(n_projects)) %>%
  slice_head(n = 25) %>%
  ggplot(aes(x = n_projects, y = reorder(country_name, n_projects),
             fill = region)) +
  geom_col(width = 0.75) +
  geom_text(aes(label = n_projects), hjust = -0.2, size = 3) +
  scale_fill_brewer(palette = "Set1", na.value = "grey70") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(title = "Top 25 Countries by Carbon Offset Project Count",
       subtitle = "Karnik et al. 2024 — 575 NbS projects across 55 countries",
       x = "Number of Projects", y = NULL, fill = "Region")
```

### 13.2 Join with WGI Data

```{r karnik-join}
df_k <- df %>%
  left_join(karnik_cty, by = "iso3c") %>%
  mutate(
    has_project  = !is.na(n_projects),
    n_projects   = replace_na(n_projects, 0),
    log_projects = log1p(n_projects)
  )

cat(sprintf("Countries with ≥1 project: %d / %d\n",
            sum(df_k$has_project), nrow(df_k)))
```

### 13.3 Correlations: Project Count × WGI

```{r karnik-cor}
cor_k <- map_dfr(wgi_vars, function(wv) {
  tmp <- df_k %>% filter(!is.na(.data[[wv]]))
  ct  <- cor.test(tmp$log_projects, tmp[[wv]],
                  method = "spearman", exact = FALSE)
  tibble(
    wgi_var   = wv,
    wgi_label = wgi_codes[wv],
    rho       = ct$estimate,
    p_value   = ct$p.value,
    sig       = case_when(p_value < 0.001 ~ "***",
                          p_value < 0.01  ~ "**",
                          p_value < 0.05  ~ "*",
                          TRUE            ~ "")
  )
})

cor_k %>%
  mutate(across(c(rho, p_value), ~ round(.x, 4))) %>%
  knitr::kable(
    col.names = c("WGI Dim.", "Label", "Spearman ρ", "p-value", "Sig."),
    caption   = "Spearman ρ: log1p(N projects) × WGI (2024)"
  )
```

```{r karnik-cor-bar, fig.height=4, fig.width=7}
cor_k %>%
  ggplot(aes(x = rho, y = reorder(wgi_label, rho), fill = rho > 0)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = paste0(round(rho, 3), sig),
                x     = rho + ifelse(rho >= 0, 0.01, -0.01)),
            hjust = ifelse(cor_k$rho >= 0, 0, 1), size = 3.8) +
  geom_vline(xintercept = 0, linewidth = 0.5) +
  scale_fill_manual(values = c("TRUE" = "#1a9850", "FALSE" = "#d73027")) +
  scale_x_continuous(limits = c(-0.5, 0.7), expand = expansion(mult = 0.02)) +
  labs(title = "Spearman ρ: Carbon Offset Projects × WGI",
       subtitle = "log1p(N projects per country); * p<0.05  ** p<0.01  *** p<0.001",
       x = "Spearman ρ", y = NULL)
```

### 13.4 Potential vs. Realised: NbS Gap Analysis

Countries with high biophysical potential but few market projects represent the
largest implementation gap. Color encodes income group; size encodes number of
market projects.

```{r karnik-gap, fig.height=6, fig.width=9}
df_k %>%
  filter(total_ncs_tco2e > 0, !is.na(PV.EST)) %>%
  mutate(
    log_ncs     = log10(total_ncs_tco2e),
    label       = if_else(
      (total_ncs_tco2e > quantile(total_ncs_tco2e, 0.85, na.rm = TRUE) &
       n_projects == 0) |
      n_projects > 20,
      adm0_name, NA_character_
    )
  ) %>%
  ggplot(aes(x = PV.EST, y = log_ncs,
             color = income_level, size = n_projects + 0.5)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = label), size = 2.6, max.overlaps = 15,
                  segment.color = "grey55", show.legend = FALSE) +
  scale_color_brewer(palette = "RdYlBu", name = "Income", na.value = "grey60") +
  scale_size_continuous(name = "Market projects", range = c(1.5, 8),
                        breaks = c(0, 5, 20, 50, 100, 200)) +
  labs(
    title    = "Biophysical NbS Potential vs. Political Stability",
    subtitle = "Point size = number of market carbon offset projects (Karnik 2024)",
    x        = "Political Stability (WGI 2024)",
    y        = "Total NbS Potential (log₁₀ tCO₂e/yr)"
  ) +
  theme_minimal(base_size = 12)
```

### 13.5 Regression: Projects ~ NbS Potential + WGI

Do countries with higher governance implement more market-based NbS, *conditional*
on biophysical potential?

```{r karnik-reg, fig.height=5}
df_kreg <- df_k %>%
  filter(total_ncs_tco2e > 0, !is.na(income_level)) %>%
  mutate(log_ncs = log10(total_ncs_tco2e),
         log_proj = log1p(n_projects))

# OLS on log1p(n_projects) — robust to the many-zeros structure
m_ols <- lm(log_proj ~ log_ncs + CC.EST + GE.EST + PV.EST +
              RL.EST + RQ.EST + VA.EST + income_level,
            data = df_kreg)

cat(sprintf("OLS adj. R² = %.3f  |  N = %d\n",
            summary(m_ols)$adj.r.squared, nobs(m_ols)))

tidy(m_ols, conf.int = TRUE) %>%
  filter(grepl("^(CC|GE|PV|RL|RQ|VA)", term)) %>%
  mutate(term = wgi_codes[term]) %>%
  ggplot(aes(x = estimate, xmin = conf.low, xmax = conf.high,
             y = reorder(term, estimate),
             color = p.value < 0.05)) +
  geom_pointrange(size = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_color_manual(values = c("TRUE" = "#d73027", "FALSE" = "grey55"),
                     guide = "none") +
  labs(
    title    = "OLS: log1p(N Carbon Offset Projects) ~ WGI",
    subtitle = sprintf("Controlling for log10(NbS potential) and income  |  Adj. R² = %.3f",
                       summary(m_ols)$adj.r.squared),
    x = "OLS Coefficient (95% CI)", y = NULL
  ) +
  theme_minimal(base_size = 12)
```

---

## 14. Export Merged Dataset

```{r export}
export_cols <- c(
  "iso3c", "adm0_name", "adm0_ha",
  nbs_biome, nbs_activity, nbs_grand,
  wgi_vars, "income_level", "region",
  "n_projects", "n_ad", "n_ifm", "n_arr"
)

df_export <- df_k %>%
  select(any_of(export_cols))

write_csv(df_export, here("data/nbs_wgi_merged.csv"))
cat("Exported", nrow(df_export), "countries →",
    here("data/nbs_wgi_merged.csv"), "\n")
```

---

## Session Info

```{r session-info}
sessionInfo()
```
